---
title: "ID33-Wilson RSV Correlates Analysis (Version 3.0)"
shorttitle: "ID33-Wilson RSV Correlates Analysis (Version 3.0)"
from:
 - name: Peter Gilbert
   email: pgilbert@fredhutch.org
   contact: yes
 - name: Youyi Fong
   email: yfong@fredhutch.org
   contact: yes
 - name: Ying Huang
   email: yhuang@fredhutch.org
   contact: no
 - name: Lars Wim Paul van der Laan 
 - name: Wenbo Zhang
 - name: Bhavesh Borate
to:
 - name: Iksung Cho
 - name: Matthew Lawlor 
 - name: Lou Fries
dropcc: false
cc:
 - name: Alicia Sato
 - name: Eva Chung
 - name: Michelle Chung
includesummary: false
summary: 
  A short summary about your report. 
bibliography: bibliography.bib
lof: true
lot: true
toc: true
header-includes:
   - \usepackage{natbib}
   - \usepackage{float}
   - \usepackage{amsmath}
   - \usepackage{comment}
   - \usepackage{subfig}
   - \usepackage[colorlinks,citecolor=red,urlcolor=blue,bookmarks=false,hypertexnames=true]{hyperref}
output: 
  VISCtemplates::visc_pdf_document: default
  VISCtemplates::visc_word_document: default
---
\counterwithin{table}{section}
\counterwithin{figure}{section}
\newcommand{\pathCoRinput}{CoR/input}
\newcommand{\pathThreshinput}{nonparam_threshold/input/y2_Vaccine}
\newcommand{\pathThreshhinput}{nonparam_threshold/input/y1_Vaccine}
\newcommand{\pathThreshpooledinput}{nonparam_threshold/input/y2_Pooled}
\newcommand{\pathThreshhpooledinput}{nonparam_threshold/input/y1_Pooled}
<!---
Program Name: RSV correlates analysis
Creation Date: 2020-04-27
Full name of author: 
Project or Protocol: RSV correlates
Purpose or description of program: Report results from RSV correlates analysis
Location of program: https://github.com/FredHutch/RSVcorrelatesAnalysis
--->

```{r package-loading-and-options, warning=FALSE, include=FALSE, message=FALSE}

library(RSVcorr)
library(vimp)

# Add additional packages needed here 
# Only works for CRAN packages, use library() for packages on Bioconductor or GiHub
packages <- c("here", "conflicted", "tidyverse", "knitr", "kableExtra", "VISCfunctions", "gridExtra", "tidyr")

# If a package is installed, it will be loaded
# If any are not, the missing package(s) will be installed and loaded
package.check <- lapply(packages, FUN = function(x) {
  if (!require(x, character.only = TRUE)) {
    # VISCfunctions must be installed in seperately (not in CRAN)
    if (x == 'VISCfunctions')
      stop('The package "VISCfunctions" must be installed through GitHub: https://github.com/FredHutch/VISCfunctions.git')
    
    install.packages(x, dependencies = TRUE,repos = "http://cran.us.r-project.org")
  }
  library(x, character.only = TRUE)
})

# Enforcing specific version of pandoc
if (numeric_version(rmarkdown::pandoc_version()) < numeric_version('2.0'))
  stop('pandoc must be at least version "2.0')

# Set some knitr options
# This will always generate two figures (pdf and png, which can be handy for presentations)
opts_chunk$set(tidy = TRUE, cache = FALSE, message = TRUE, warning = TRUE,
               echo = FALSE, dev = c("png", "pdf"), dpi = 200,
               out.width = '100%', out.extra = '', fig.align = "center",
               fig.pos = "H", fig.height = 7, fig.width = 7, fig.path = "C:/figure-latex/")

# NA's will be blank in tables
options(knitr.kable.NA = '')

# I use Claus Wilke's colorblindr package for the experiment colors. 
# devtools::install_github("clauswilke/colorblindr")

# spelling::spell_check_files(here::here("correlates_analysis", "correlates_analysis.pdf"))

# # functions for this report
# source(here::here("R", "functions.R"))
```

```{r conflicted-preferences, message=FALSE}

# filter() is used in both dplyr and stats, so need to set the preference to dplyr
conflict_prefer("filter", "dplyr")
conflict_prefer("summarise", "dplyr")
```

```{r word-pdf-functions}

# Need to see if PDF or Word Output
current_output_type <- knitr::opts_knit$get('rmarkdown.pandoc.to')

# If interactive setting to pandoc for easier visualization
output_type <- ifelse(
  !is.null(current_output_type) && current_output_type == 'latex', 
  'latex', 
  'pandoc'
  )
kable_warnings <- ifelse(output_type == 'latex', TRUE, FALSE)
pandoc_markup <- ifelse(output_type == 'pandoc', TRUE, FALSE)

# Setting up references depending on Word or PDf output
visc_ref <- function(ref_in) {
  ifelse(knitr::opts_knit$get('rmarkdown.pandoc.to') == 'latex', 
          paste0('\\ref{', ref_in, '}'), 
          paste0('\\@ref(', ref_in, ')')) 
}

visc_clearpage <- function() {
  ifelse(knitr::opts_knit$get('rmarkdown.pandoc.to') == 'latex', 
          '\\clearpage', 
         '#####')
} 
```


`r visc_clearpage()`

\textbf{Version 2.0 of the RSV immune correlates report} was updated with analyses requested by Dr. Keith Klugman: Section [Posthoc analysis to assess fold change from maternal baseline to cord blood in immunologic biomarkers as CoR for endpoint 2].

\textbf{Version 3.0 of the RSV immune correlates report} contains enhancements to the non-parametric threshold-risk plots: Figures \ref{fig:PLOT_EIA_log10d14_y1_Vaccine_pointwiseCI}--\ref{fig:PLOT_RSVB_log10cord_y2_Pooled_pointwiseCI} have been revamped with point-wise confidence intervals, superimposed marker-specific reverse CDF (RCDF) plots, and the threshold of no more events has been marked by a vertical red line.


# Executive Summary

```{r, child=c('ExecutiveSummary.Rmd')}
```

# Materials and methods

The materials and methods are fully described in the Statistical Analysis Plan document (Section [Statistical Analysis Plan]). Figure \ref{fig:endpoints_venn_diagrams} shows how the endpoints primarily evaluated for correlates analyses in this report (endpoints 1 and 2) relate to the endpoints defined in the efficacy analyses. 

\begin{figure}[H]
    \centering
    \includegraphics[width=.8\textwidth]{endpoints_venn_diagrams}
    \caption{Venn diagram showing relationship of endpoints 1 and 2 that were evaluated for correlates analyses with endpoints defined in the efficacy analyses. Endpoint 3, RSV LRTI with Severe Hypoxemia without cough, is closely related to and a subset of endpoint 2.}
    \label{fig:endpoints_venn_diagrams}
\end{figure}


# Risk Score analysis

```{r, child=c('SL_risk_score/SL_risk_score.Rmd')}
```


`r visc_clearpage()`

# Immune biomarkers unsupervised exploratory analysis (preparation for correlates analyses)

```{r, child=c('exploratory/exploratory_plots.Rmd')}
```


`r visc_clearpage()`

# Correlates of risk

<!-- \pathCoRinput -->

```{r, child=c('CoR_logistic_regression/CoR_obj1.Rmd')}
```

```{r, child=c('CoP_nonparametric_threshold/estimated_conditional_risk.Rmd')}
```

```{r, child=c('CoR_logistic_regression/CoR_obj2.Rmd')}
```


```{r, child=c('objective3/SL_estimated_optimal_surrogate.Rmd')}
```


```{r, child=c('CoR_logistic_regression/CoR_posthoc.Rmd')}
```




`r visc_clearpage()`

# Correlates of protection

```{r, child=c('CoP_logistic_regression/CoP_baselineVE.Rmd')}
```

```{r, child=c('CoP_principal_stratification/CoP_Ying.Rmd')}
```


```{r, child=c('CoP_logistic_regression/CoP_mediator.Rmd')}
```

```{r, child=c('CoP_logistic_regression/CoP_posthoc.Rmd')}
```



`r visc_clearpage()`

# Appendix

## Risk score analysis

```{r, child=c('SL_risk_score/SL_risk_score_Appendix.Rmd')}
```

## Superlearner regression for calculating the estimated optimal surrogate

```{r, child=c('objective3/SL_estimated_optimal_surrogate_Appendix.Rmd')}
```


`r visc_clearpage()`

```{r, child=c('Keith_analysis.Rmd')}
```

# References

<div id="refs"></div>

`r visc_clearpage()`

# Reproducibility information

```{r Software-Session-Information, results='asis', message=FALSE, warning=kable_warnings}
### Chunk to create reproducibility tables
# loading in rmarkdown and VISCtemplates so we can capture version number
if (any(installed.packages()[,1] == 'rmarkdown')) suppressWarnings(library(rmarkdown))
if (any(installed.packages()[,1] == 'VISCtemplates')) suppressWarnings(library(VISCtemplates))

my_session_info <- VISCfunctions::get_session_info()
kable(
  my_session_info$platform_table,
  format = output_type,
  booktabs = TRUE,
  linesep = '',
  caption = "Reproducibility software session information"
  ) %>%
  kable_styling(font_size = 8, latex_options = "hold_position")
```

```{r Software-Package-Version-Information, results='asis', warning=kable_warnings}
kable(
  my_session_info$packages_table,
  format = output_type, booktabs = TRUE,
  linesep = '',
  caption = "Reproducibility software package version information"
  ) %>%
  kable_styling(font_size = 8, latex_options = "hold_position")
```


# Statistical Analysis Plan
